---
title: "Mini-Project 01"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

# Data Visualization Project 01

_revised version of mini-project 01 goes here_


```{r setup, include=FALSE, message=FALSE, warnings=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
```


```{r}
library(tidyverse)
library(magrittr)
library(dplyr)
library(readr)
library(stringr)
```

# Data Exploaration

```{r cars}
fuel_link <- "../data/fuel.csv"
fuel <- read_csv(fuel_link)
glimpse(fuel);head(fuel)
```


```{r fig.align="center", echo = FALSE,fig.width = 10}
# Example of disproportionate Data. Natural gas penetration is extremely low compared to other sources
# of fuel.
p <- ggplot(data = fuel,
            mapping = aes(x = fuel_type_1, fill = fuel_type_1))
p + geom_bar() + coord_flip()

```



## Few Points regarding the "Fuel" Data set
1. There are too many columns some of them are not used at all.
2. Many missing values in the form of NAs or zeros in some cases.
3. The data set combines 3 main fuel types; Gas, Electric and Hybrid
4. FT stands for Fuel type. Fuel Type 2 is meant for vehicles that have dual cell but this is not
   indicated in the data set and hence it will be ignored.
5. Depending on the model of the vehicle data may not exist since 1986.
6. Other relevant and important variables are not included; Weight, Horsepower Rating and starting MSRP of 
   vehicle.
7. Due to disproportionate data, some entries will be ignored such as the the use of Natural Gas. See the      bar chart above.
8. Data prior to 2000 since life expectancy of a car is around 12 - 15 years according 


# Data Preparation
1. To prepare for Visualization, a subset of the data will be selected 
2. A new column called "car_type" will be created based on the "class" column
3. A new column will be created called "trs_type" based on transmission type
4. A new column will be created called "energy_type" based on "c" column
5. Outlines will be ignored.



## Selecting a subset of the data beased on model year and selected columns
```{r}
fuel_main <- fuel  %>% 
  select("year","make", "model","class", "drive", "transmission", "transmission_type",
         "engine_cylinders","engine_displacement", ,"fuel_type","fuel_type_1","city_mpg_ft1",
         "highway_mpg_ft1","combined_mpg_ft1", "save_or_spend_5_year", "range_ft1") %>%
  filter(year > 1995)

```



## Checking the existenc of "NA" in each column
```{r   message=FALSE, warnings=FALSE}
fuel_na <- fuel_main %>%
  summarise_all(funs(sum(is.na(.))))
table(fuel_na)
fuel_na
```

## Most Common values of class
```{r fig.align="center", echo = FALSE,fig.width = 10}
p1 <- ggplot(data = fuel_main,
            mapping = aes(x=reorder(class, class, function(x)+length(x)), fill = class))
p1 + geom_bar() + coord_flip() + guides(fill = "none")

```



```{r}
fuel_sel <- fuel_main  %>%
  mutate(car_type = case_when(class == "Compact Cars" | class == "Sport Utility Vehicle - 2WD"
                               | class == "Small Station Wagons" ~ "Small_Family",
                              class == "Midsize Cars" | class == "Sport Utility Vehicle - 4WD"
                               | class == "Large Cars" ~ "Mid_Family",
                              class == "Standard Pickup Trucks 2WD" ~ "Work",
                              class == "Subcompact Cars" & class > 3.0 ~ "High_PRF",
                              class == "Subcompact Cars" & class <= 3.0 ~ "Small_Family",
                              class == "Two Seaters" & class > 3.0 ~ "High_PRF",
                              class == "Two Seaters" & engine_displacement <= 3.0 ~ "Small_Family",
                              class == "Minicompact Cars" & engine_displacement > 3.0 ~ "High_PRF",
                              class == "Minicompact Cars" & engine_displacement <= 3.0 ~ "Small_Family",
                              .default = "Other"))%>%
  mutate(trs_type = case_when(str_detect(tolower(transmission),"auto") ~ "Automatic",
                              str_detect(tolower(transmission),"man") ~ "Manual",
                              str_detect(tolower(transmission),"na") ~ "Automatic")) %>%
  mutate(energy_type = case_when(str_detect(tolower(fuel_type_1),"gas") ~ "Fossil",
                              str_detect(tolower(fuel_type_1),"diesel") ~ "Fossil",
                              str_detect(tolower(fuel_type_1),"electricity") ~ "Renewable"))

```


```{r}
fuel_sel <- fuel_sel[fuel_sel$car_type != "Other", ]
fuel_sel <- fuel_sel[complete.cases(fuel_sel$trs_type),]
```

### Data Verification after Preparation

```{r}
dim(fuel);dim(fuel_main);dim(fuel_sel)
```


```{r}
pc <- ggplot(data = fuel_sel,mapping = aes(x = trs_type, fill = trs_type))
pc + geom_bar(na.rm = TRUE) + coord_flip()

```


# Visualizations

You can also embed plots, for example:
to try to declutter the plot I will be creating a new variable dip_mean = mean of engine displacement for each group of cylinders.

```{r}
fuel_sel_cyl <- fuel_sel %>%
  group_by(engine_cylinders) %>%
  mutate(disp_mean = mean(engine_displacement))
```


```{r warnings=FALSE,  fig.align="center", echo = FALSE,fig.width = 10, fig.height=8}
       
mpg_p <-ggplot(data = fuel_sel_cyl[!(is.na(fuel_sel_cyl$trs_type)),], filter (fuel_sel_cyl,energy_type %in% c("Fossil")),
           mapping = aes(x = disp_mean,
                          y = combined_mpg_ft1,
                          color = car_type)) + 
  geom_point(size = 3) +
  #scale_size(guide = "none") +
  geom_smooth(method = "loess") +
  #scale_x_binned() +
  ylim(5,55) + 
  facet_wrap( ~ trs_type, ncol = 1) +
  labs(
    x = "Engine Size in Liters",
    y = "Combined Miles per Gallons",
    title = "Miles per Gallon vs Engine Displacment",
    caption = "Data: 1984-2017 fuel economy data produced during vehicle testing
                  at the (EPA) National Vehicle and Fuel Emissions Laboratory."
  ) + 
  guides(color=guide_legend(title="Type of Car"))+ 
  theme_bw()
print_mpg <- suppressWarnings(print(mpg_p))
```



```{r warnings=FALSE,  fig.align="center", echo = FALSE,fig.width = 9}
drive_p <-ggplot(data = fuel_sel[!(is.na(fuel_sel$trs_type)),], filter (fuel_sel,energy_type %in% c("electricity")),
           mapping = aes(x = round(year,0),
                          y = range_ft1,
                          color = car_type)) + 
  geom_point(aes(size=drive)) +
  scale_size_discrete(guide = "none") +
  geom_smooth(method = "loess")+
  labs(
    x = "Time in Years",
    y = "Range in Miles on Single Charge",
    title = "Timeline of EV Vehicle Range per Car Type",
    subtitle = "Color and Size: Car Type and Driveterain",
    caption = "Data: 1984-2017 fuel economy data produced during vehicle testing
                  at the (EPA) National Vehicle and Fuel Emissions Laboratory.")
print_drive <- suppressWarnings(print(drive_p))


```


```{r fig.align="center", echo = FALSE,fig.width = 12}

fuel_p <- ggplot(data = fuel_sel,
            mapping = aes(x = year, y = round(save_or_spend_5_year/1000,0), fill = car_type))
fuel_p + geom_col() + 
  facet_wrap( ~ energy_type, ncol = 1, scales = "free")+
  labs(x = NULL, y = "Average Money Saved per Car Type",
       title = "Renewable vs Fossil Fuel Expected Savings Timeline",
       subtitle = "Colored by Car Type",
       caption = "Data: 1984-2017 fuel economy data produced during vehicle testing
                  at the (EPA) National Vehicle and Fuel Emissions Laboratory.")
```


