---
title: "Mini-Project 02"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

# Data Visualization Project 02

_revised version of mini-project 02 goes here_


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction:
In the project I will be working on two data sets:
- The Florida lakes Data set for Spatial Visualization
- The WestRoxbury data for Interactive and model based plotting

Importing All required Libraries:
```{r include=FALSE}
library(tidyverse)
library(gapminder)
library(scales)
library(dplyr)
library(socviz)
library(ggrepel)
library(dichromat)
library(RColorBrewer)
library(ggthemes)
library(cowplot )
library(hrbrthemes)
library(ggridges)
library(maps)
library(sf)
library(plotly)
library(sp)
library(reshape2)
library(corrplot)
library(flexdashboard)
```


# Data Exploration and Summary

## Florida Lakes Data Set

### Reading the data
```{r}

loc_lakes <- "../data/Florida_Lakes.shp"
## Read the shapefile into R

fl_lakes <- st_read(loc_lakes,quiet = TRUE)
```

### Data Summary - Florida Lakes:

```{r warnings=FALSE}
head(fl_lakes)
summary(fl_lakes)
lakes_na <- fl_lakes %>%
  as.data.frame() %>%
  summarise_all(funs(sum(is.na(.))))
lakes_na
```


### Finding the top Counties with highest number of lakes and the average size of lakes 

```{r warnings=FALSE}
# finding the to 10 largest lakes in all Florida lakes
top_10_lakes <- fl_lakes %>% 
  as_data_frame() %>% 
  arrange(desc(SHAPEAREA)) %>%
  top_n(SHAPEAREA, n = 10)
# create an sf object from the top ten data frame
top_10_lakes_sf <- st_as_sf(top_10_lakes)
```


```{r}
# I need to space out "East Lake Tohopekaliga" since its too close to "Lake Tohopekaliga" by assigning it
# a different x nudge
nudges <- top_10_lakes_sf %>%
  mutate(my_nudge_x = ifelse(NAME=='East Lake Tohopekaliga',0.5,-0.75))
```


## Part I: Spatial Visualization

I will plot all florida lakes highlighting the top 10 largest lakes on the map
```{r warnings=FALSE,  fig.align="center", echo = FALSE,fig.width = 20, fig.height=16}

fl_lakes %>%
  ggplot() +
  geom_sf(col = "salmon3",fill = "white") +
  geom_sf(data = top_10_lakes_sf, aes(fill = log(SHAPEAREA))) +
  scale_fill_viridis_c(option = "plasma") +
  geom_sf_label(data = top_10_lakes_sf,
               aes(label = NAME), fill = "white", col = "grey20",
               size = 6,nudge_x = nudges$my_nudge_x, family = "serif") +
  labs(title = "Top 10 Largest Lakes in Florida") +
  theme(plot.title = element_text(size = 20, face = "bold", colour = "grey30"))
```
`

## West Roxbury Data Set
### Reading the data


```{r warnings=FALSE}
data_url <- "../data/WestRoxbury.csv"
west_rox <- read.csv(data_url)
```


### Data Summary - WestRoxbury

```{r}
summary(west_rox)
head(west_rox)
west_rox$REMODEL <- as.factor(west_rox$REMODEL)

```

```{r}
#checking if there are any NA values present in the data set
rox_na <- west_rox %>%
  summarise_all(funs(sum(is.na(.))))
rox_na
#No NA values exist in the data set
```


### Exploring the data

The following two chunks of code is to explore the distributions of the columns 
excluding the remodeling attribute since its a character
```{r}
west_rox_mod <- west_rox[!(colnames(west_rox) %in% "REMODEL")] # removing the REMODEL column
west_long <- west_rox_mod %>%                         
  pivot_longer(colnames(west_rox_mod)) %>% 
  as.data.frame()

  
```

### Attribute distrbutions
```{r message=FALSE, warning = FALSE}
ggp1 <- ggplot(west_long, aes(x = value)) +    # Draw each attribute
  geom_histogram() + 
  facet_wrap(~ name, scales = "free")
ggp1
```
- disregarding the aesthetics of the graph, we can see that there seems to be a correlation between: living area (and hence Gross Area), 
Tax and Value of the house. We will explore that below under correlation
- Data for many distributions is not balanced.

### Correlation table:

```{r message=FALSE, warning = FALSE}
corrplot(cor(west_rox_mod))
```
Few takeaways from the correlation graph:
1. Strong correlation between House value, Tax, Area, Gross area, Rooms and Bedrooms.
2. Only one attribute of any pairs (or more) of strongly correlated attributes will be selected for prediction. 
3. Since gross area encompasses living area using one of them in the model will be enough
4. Kitchen has almost no impact.
5. The age of the house has a slight negative impact on the value of the house.


### Building a model to predict the value of the house
```{r message=FALSE, warning = FALSE}
library(GGally)
library(broom)
west_glm <- glm(TOTAL.VALUE ~ GROSS.AREA + ROOMS + FLOORS,
                     data = west_rox)
hous_coefs <- tidy(west_glm, conf.int = TRUE) %>%
  filter(term != "(Intercept)")
```

### Building test data for prediction
```{r}
hous_new <- expand.grid(GROSS.AREA = seq(800, 8000, by = 100), # area start around 800 ft to roughly 8000
                   ROOMS = seq(1,12 , by = 1), # from the summary number of rooms between 1 & 12
                   FLOORS = seq(1, 3, by = 1)) # max. number of floor is 3 from summary
head(hous_new)
```

FLOORS = seq(0, 3, by = .1))
### Using the model to make prediction
```{r}
glm_pred <- augment(west_glm, newdata = hous_new, se_fit = TRUE)
head(glm_pred)
```

```{r}
# The below extra data frames are created to be able to add a rectangle and text on one of the facet plots not all
ann_rect <-  pred_hous_value[!(colnames(pred_hous_value) %in% ".se.fit")] # This for the rectangle

ann_text <- data.frame(GROSS.AREA = 7000,.fitted = 475,lab = "Text", # This one for the label.
                       FLOORS = factor(1,levels = c("1","2","3")))
```


## Part II: Model based Visualization
```{r warnings=FALSE,  fig.align="center", echo = FALSE,fig.width = 12, fig.height=8}
ggplot(glm_pred,
       aes(x = GROSS.AREA, y = .fitted)) +
  geom_point(data = glm_pred, aes(col = ROOMS), alpha = 0.5) +
  geom_smooth(method = "loess", color = "black") +
  facet_wrap(vars(FLOORS)) + 
  scale_colour_viridis_c() +
  theme_minimal() +
  geom_rect(data = ann_rect %>%
              filter(FLOORS == 1),
            aes(xmin = 4000, xmax = 5200, ymin = 420, ymax = 520), alpha = .01, fill="plum", inherit.aes = FALSE)+
  geom_text(data = ann_text,label = "Vertical variations\n due to # floors", hjust = 0.5) +
  labs(title = "Predicted House Values as a Function of Total Area, Floors and Number of Rooms") +
  theme(plot.title = element_text(size = 12, face = "bold", colour = "dodgerblue4"))
  
```


# Part III : Interactive Visualization
This part will be completed using a separate markdown file using FlexDashboard template.
